---
title: "Illustration of mix-SQP solver applied to a small data set, and a large one"
author: "Youngseok Kim and Peter Carbonetto"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mixSQP-intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE,comment = "#",fig.width = 5,
                      fig.height = 3,fig.align = "center",
                      fig.cap = "&nbsp;",dpi = 120)
```

This vignette illustrates mix-SQP, a fast algorithm for
maximum-likelihood estimation of mixture proportions, and compares its
performance against an interior-point method.

## Environment set-up

We begin by loading the mixSQP package.

```{r load-pkgs}
library("mixSQP")
```

Next, initialize the sequence of pseudorandom numbers.

```{r set-seed}
set.seed(1)
```

## Generate a small, simulated data set

Let's begin with a smaller example with 50,000 samples.

```{r sim-data-small}
sim <- simulatemixdata(5e4,20)
```
