---
title: "Illustration of mixsqp applied to a small data set, and a large one"
author: "Youngseok Kim, Peter Carbonetto and Matthew Stephens"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_document:
    theme: readable
    before_body: header.html
    after_body: footer.html
vignette: >
  %\VignetteIndexEntry{mixsqp-intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE,results = "hold",comment = "#")
```

In this vignette, we illustrate the use of the sequential quadratic
programming (SQP) algorithm implemented in `mixsqp`, and compare its runtime and accuracy
against an interior-point (IP) solver implemented by the
[MOSEK][mosek] commercial software (called by the "KWDual"
function in the [REBayes package][rebayes]).

## Environment set-up

Load the `mixsqp` and `Rmosek` packages. (The Rmosek package is only
used to check the version of MOSEK.)

```{r load-pkgs, message=FALSE}
library(mixsqp)
library(Rmosek)
```

Next, initialize the sequence of pseudorandom numbers.

```{r set-seed}
set.seed(1)
```

## Generate a small data set

We begin with a small example to show how `mixsqp` works.

```{r sim-data-small}
L <- simulatemixdata(1000,20)$L
dim(L)
```

This call to `simulatemixdata` created an $n \times m$ conditional
likelihood matrix for a mixture of zero-centered normals, with $n =
1000$ and $m = 20$. By default, `simulatemixdata` normalizes the rows
of the likelihood matrix so that the maximum entry in each row is 1.

## Fit mixture model

Now we fit the mixture model using the SQP algorithm:

```{r fit-model-mixsqp-small}
fit.sqp <- mixsqp(L)
```

In this example, the SQP algorithm converged to a solution in a small
number of iterations.

By default, `mixsqp` outputs information on its progress. It begins by
summarizing the optimization problem and the algorithm settings used.
(Since we did not change these settings in the `mixsqp` call, all the
settings shown here are the default settings.)

After that, it outputs, at each iteration, information about the
current solution, such as the value of the objective ("objective") and
the number of nonzeros ("nnz").

The "max(rdual)" column shows the quantity used to assess convergence.
It reports the maximum value of the "dual residual"; the SQP solver
terminates when the maximum dual residual is less than `conv.tol`,
which by default is $10^{-8}$. In this example, we see that the dual
residual shrinks rapidly toward zero.

Another useful indicator of convergence is the "max.diff" column---it
reports the maximum difference between the solution estimates at two
successive iterations. We normally expect these differences to shrink
as we approach the solution, which is precisely what we see in this
example.

This information is also provided in the return value, which we can
use, for example, to create a plot of the objective value at each
iteration of the SQP algorithm:

```{r plot-sqp-progress, fig.height=5,fig.width=7, fig.align="center"}
numiter <- nrow(fit.sqp$progress)
plot(1:numiter,fit.sqp$progress$objective,type = "b",
     pch = 20,lwd = 2,xlab = "SQP iteration",
     ylab = "objective",xaxp = c(1,numiter,numiter - 1))
```

To assess the accuracy of the SQP solution, we use the IP solution as
a point of comparison:

```{r fit-model-kwdual-small}
fit.ip <- mixkwdual(L)
```

In this example, the IP and SQP solutions achieve nearly the same
objective value:

```{r compare-solutions-smaller}
cat(sprintf("Objective at SQP solution: %0.16f\n",fit.sqp$value,digits = 16))
cat(sprintf("Objective at IP solution:  %0.16f\n",fit.ip$value,digits = 16))
cat(sprintf("Difference in objectives:  %0.4e\n",fit.sqp$value - fit.ip$value))
```

## Comparing the SQP and IP solvers in a large data set

We observed that the SQP and IP methods achieve nearly the same
solution quality in the example above. Here, we explore the
computational properties of the SQP and IP algorithms in a larger data
set.

As before, we compute the $n \times m$ likelihood matrix for a mixture
of zero-centered normals. This time, we use a finer grid of $m = 100$
normal densities, as well as many more samples.

```{r sim-data-large}
L <- simulatemixdata(1e5,100)$L
dim(L)
```

Now we fit the model using the SQP algorithm,

```{r fit-model-sqp-large}
timing <- system.time(fit.sqp <- mixsqp(L))
cat(sprintf("Computation took %0.2f seconds\n",timing["elapsed"]))
```

and using the IP method,

```{r fit-model-ip-large}
timing <- system.time(fit.ip  <- mixkwdual(L))
cat(sprintf("Computation took %0.2f seconds\n",timing["elapsed"]))
```

The SQP algorithm was considerably faster than the IP solver, and it
converged to a solution with nearly the same objective value as the IP
solution:

```{r compare-solutions-large}
cat(sprintf("Objective at SQP solution: %0.16f\n",fit.sqp$value,digits = 16))
cat(sprintf("Objective at IP solution:  %0.16f\n",fit.ip$value,digits = 16))
cat(sprintf("Difference in objectives:  %0.4e\n",fit.sqp$value - fit.ip$value))
```

## Session information

This next code chunk gives information about the computing environment
used to generate the results contained in this vignette, including the
version of R and the packages used.

```{r session-info}
sessionInfo()
```

Since we used `Rmosek`, it is also helpful to record the version of
MOSEK that was used.

```{r mosek-info}
mosek_version()
```

[mosek]: https://www.mosek.com
[rebayes]: https://cran.r-project.org/package=REBayes
